Microbiome and hypertension

### 2018-12-20 rrbiome.Rmd
PCA-plots

library(dplyr)
library(tibble)
library(phyloseq)
library(microbiome)
library(finriskmetagcommon)
library(knitr)
library(tidyr)
library(vegan)

@@

if (!exists('phfinrisk_metadatadesc')) {
  # this is a hand-compiled data.frame that contains metadata information for variables utilized in our analyses
  data("phfinrisk_metadatadesc", package="finriskmetagcommon")
}
names.dset <- filter(phfinrisk_metadatadesc, Ignored.Covariate.in.Cross.Sectional.Analysis.Aaro20181115==0)

@@

```{r alpha diversity scatterplot, fig.width=8, fig.height=4}

ggplot(dset.alpha, aes(x=diversities_shannon, y=MAP)) +
    geom_point(size=1, shape=1) +
    coord_fixed() +
    ggplot2::theme_minimal() 
```

@@

Matriisin teko _bc_dissimilarity_matrix.R ja käyttö comp_adonis_covariates.R

@@

@@

@@

attr(bray.dist.m, "method")
dist.matrix <- as.matrix(bray.dist.m)
attr(dist.matrix, "method") <- "bray"

@@

```{r beta heatmap, eval = FALSE}

ggplot(data = dist.matrix, aes(x = metrics, y = teams)) +
      geom_tile(aes(fill = performance))

@@

>>>>

library(dplyr)
library(tibble)
library(phyloseq)
library(microbiome)
library(finriskmetagcommon)
library(knitr)
library(tidyr)
library(vegan)
library(reshape)
library(ggpmisc)

@@

data("phfinrisk_metadatadesc", package="finriskmetagcommon")
names.dset <- filter(phfinrisk_metadatadesc,
                     Ignored.Covariate.in.Cross.Sectional.Analysis.Aaro20181115==0)

@@

```{r alpha diversity scatterplot, fig.width=10, fig.height=4}

ggplot((dset.alpha %>%
        select(BL_AGE, BMI, SYSTM, DIASM, PULSEPRESSURE, MAP, SEX, diversities_shannon) %>%
        gather(Covariate, Value, -SEX, -diversities_shannon)), 
       aes(x=diversities_shannon, y=Value)) +
    facet_wrap(~Covariate, ncol=4, scales = "free") +
    geom_point(size=0.1, shape=1, aes(colour = SEX)) +
    geom_smooth(method=lm, se=TRUE, size=0.5, color = 'Black') +
    stat_poly_eq(formula = "y ~ x",
                 aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")),
                 label.y = 0,
                 parse = TRUE,
                 size = 3) +   
    ylim(0, 200) +
    ggplot2::theme_minimal()

@@

Beetadiversiteetti-matriisin lasku. Tiedosista comp_bc_dissimilarity_matrix.R ja käyttö comp_adonis_covariates.R.

@@

dist.matrix <- as.matrix(bray.dist.m)
attr(dist.matrix, "method") <- "bray"

@@

```{r beta hclust, fig.width = 5, fig.height = 4, dpi = 300}
gplots::heatmap.2(dist.matrix,
                  distfun = dist,
                  trace = "none",
                  hclustfun = hclust,
                  key.title = NA,
                  dendrogram = "column",
                  density.info = "none",
                  labRow = NA,
                  labCol = NA)
```

Mallin lasku Adoniksella

@@

@@

# PCA

```{r pca}
pseq.clr <- microbiome::transform(pseq_filtered, "clr")
X <- as.matrix(t(abundances(pseq.clr)))
prcomp <- stats::prcomp(X)
```

```{r pca labels}
percentage <- round(prcomp$sdev/sum(prcomp$sdev) * 100, 2)
pca.labels <- paste0(colnames(prcomp$x), " (", paste0(as.character(percentage), "%", ")"))
```

```{r PCA plots categorical, fig.width = 10, fig.height = 10, dpi = 300}
ggplot((cbind(meta(pseq_filtered), prcomp$x[,1:2]) %>%
        select(PC1, PC2, BP_TREAT, CURR_SMOKE, DIAB, BL_USE_RX_C09, Q57X,
               BL_USE_RX_C03, BL_USE_RX_C07, BL_USE_RX_C08, HYPERTENSION, SEX) %>%
        gather(Covariate, Value, -PC1, -PC2)), 
       aes(x=PC1, y=PC2)) +
    facet_wrap(~Covariate, ncol=2, scales = "free") +
    geom_point(aes(color = Value), size = 0.5) +
    scale_color_brewer(palette="Dark2") +
    xlab(pca.labels[1]) +
    ylab(pca.labels[2]) +
    ggplot2::theme_minimal()

@@

```{r PCA plots continuous, fig.width = 10, fig.height = 10, dpi = 300}

ggplot((cbind(meta(pseq_filtered), prcomp$x[,1:2]) %>%
        select(PC1, PC2, BL_AGE, SYSTM, DIASM, BMI, MAP, PULSEPRESSURE) %>%
        gather(Covariate, Value, -PC1, -PC2)), 
       aes(x=PC1, y=PC2)) +
    facet_wrap(~Covariate, ncol=2, scales = "free") +
    geom_point(aes(color = Value), size = 0.5) +
    scale_color_gradientn(colours = rainbow(12)) +
    xlab(pca.labels[1]) +
    ylab(pca.labels[2]) +
    ggplot2::theme_minimal()
